Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies

### 2021-05-14 2021/emnlp/auto_cl_mdmt.tex

% Our method
MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model. We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$. We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$. After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}. The evolution of $\psi$ is thus defined by:
\begin{align*}
\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
\end{align*}
\begingroup
\allowdisplaybreaks
where reward $R(d)$ is computed as:
\begin{align*}
  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
\end{align*}
and where we also define:
\begin{equation}
\begin{array}{rcl}
\theta_{t+i} &=& Update\big(\theta_{t+i-1},[x^i_j,y^i_j]_{j=1}^N\big) \\ \nonumber
x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
\end{array}
\end{equation}
\endgroup
% \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.

>>>>

% Our method
MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model. We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$. We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$. After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}. The evolution of $\psi$ is thus defined by:
\begin{align*}
\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
\end{align*}
\begingroup
\allowdisplaybreaks
where reward $R(d)$ is computed as:
\begin{align*}
  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
\end{align*}
and where we also define:
\begin{equation}
\begin{array}{rcl}
\theta_{t+i} &=& Update\big(\theta_{t+i-1},[x^i_j,y^i_j]_{j=1}^N\big) \\ \nonumber
x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
\end{array}
\end{equation}
\endgroup
% \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.
